---
title: "Working with directories and lists of files"
# description: |
#   A short description of the post.
date: 2018-09-28
output:
  distill::distill_article:
    self_contained: false
    toc: TRUE
    toc_depth: 2
    toc_float: TRUE
draft: FALSE
categories: [tools]
editor_options: 
  markdown: 
    wrap: 80
---

```{r}
if (!require(pacman)){install.packages('pacman')}
p_load(tidyverse)
```

A few useful R code chunks to execute common tasks when I compile a complex research project with several sub-folders organizing the content. A very common goal is to load in the environment all the files stored in a certain sub-directory of the project.

# 1) Using `{base}dir` + `{base}load`(wrapped in `{base}lapply`) to LOAD several files

For example `base::dir` produces a character vector of the names of
files or directories in the named dir:

> `dir(path = ".", pattern = NULL, all.files = FALSE,full.names = FALSE, recursive = FALSE, > ignore.case = FALSE, include.dirs = FALSE, no.. = FALSE)`

... and I can use it to obtain a list of files which I then `base::readRDS`
using `base::lapply`:

```{r}
# Get the list of FILES 
file_names  <-  as.list(dir(
	path ="./_posts/2018-09-28-dirs-files/files", # no final {/}
	pattern = "*.Rds",
	full.names = TRUE ))
file_names

# read them in the environment
lapply( file_names, readRDS  )
```

 
# 2) Selectively load the list of FILES *ending* with ""

Similarly, I can use `base::list.files`, with the addition of the
argument `pattern = <some regex>` to screen certain files... again
followed by `base::lapply(load)`

> `list.files(path = ".", pattern = NULL, all.files = FALSE, full.names = FALSE, recursive = FALSE, ignore.case = FALSE, include.dirs = FALSE, no.. = FALSE)`


```{r}
# List of files ending with <.Rds>
file_names2 <- list.files(path = "./_posts/2018-09-28-dirs-files/files",
							 pattern = ".Rds$",
                      full.names = T # "xx$" = "ends with  xx"
                      # all.files = FALSE,
                      # full.names = FALSE, recursive = FALSE,
                      # ignore.case = FALSE, include.dirs = FALSE, no.. = FALSE)
                      )   
# read them in the environment
lapply( file_names2, readRDS  )
```

# 3) Now, a variation on the theme from tidyverse package `purrr` to EXECUTE several R scripts

Again, I use `base::list.files`, but combined with `purrr::walk(source)` - which is a more sophisticated function to loop through things and execute something.

Plus, instead of `load` I use `source` because I intend to EXECUTE several R scripts contained in a project's sub folder.

> (\*) function `purrr:walk` is specifically used for functions that don't produce an output (as opposed to `purrr:map`)


```{r}
# Get the list of R SCRIPTS [!!!]
file_names3 <- list.files(
	path = "./_posts/2018-09-28-dirs-files/files",
	pattern = ".R$",
	all.files = FALSE,  # def (= only visible)
	full.names = TRUE,  # I NEED dir name prepended
	recursive = FALSE,  # def  (= no inside sub-dir )
	ignore.case = TRUE, # (= pattern-matching be case-insensitive)
	include.dirs = FALSE, # def (subdirectory names NOT be included in recursive listings)
	no.. = FALSE) %>% # def (both "." and ".." be excluded also from non-recursive listings) 
	sort(decreasing = FALSE
		  )  
file_names3

# Execute them them in the environment
purrr::walk(.x = file_names3,
				.f = source, 
				local = FALSE, 
				echo = TRUE, 
				verbose = TRUE) 
```

 
# Reference

Check out this [blog post by Bob Rudis](https://rud.is/b/2016/07/26/use-quick-formula-functions-in-purrrmap-base-vs-tidtyverse-idiom-comparisonsexamples/) particularly interesting for comparison b/w `purrr::map` and `base` 
