---
title: R - working with lists
# description: |
#   A short description of the post.
# author:
#   - name: Nora Jones 
#     url: https://example.com/norajones
#     affiliation: Spacely Sprockets
#     affiliation_url: https://example.com/spacelysprokets
date: '2022-04-25'
output:
  distill::distill_article:
    self_contained: false
    toc: TRUE
    toc_depth: 2
    toc_float: TRUE
draft: false
categories: [tools]
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(eval = TRUE,
                      echo = TRUE, # display the source code
                      results = "hide",
                      warning = FALSE,
                      message = FALSE,
                      # fig.path = "figs/",
                      # fig.width = 7.252,
                      # fig.height = 4,
                      comment = "#>",
                      fig.retina = 3 #,
                      #R.options = list(width = 60)
)

if (!require(pacman)) install.packages("pacman")
p_load(tidyverse, 
       data.table)

```

# Theory Key Sources 

+ [Jenny Bryan's STAT545 course](https://stat545.com/r-objects.html)
+ [Vincenzo Coia'sSTAT545 course - new version](https://stat545.stat.ubc.ca/course/)
+ [Tidyr Blog](https://tidyr.tidyverse.org/reference/hoist.html)
+ [Capter 9 Introduction to Programming with R](https://discdown.org/rprogramming/lists.html#lists-list-attributes)

# [CASE 1] S.O. problem "nested-list-to-wide-dataframe"

Trying to answer to SO questions: 
https://stackoverflow.com/questions/51020513/r-nested-list-to-wide-dataframe

```{r reprex, results='markup'}
# Very nested example taken from stackoverflow.com  
x <- list()
x[[1]] <- list(uuid = "123",
               relationships = list(websites = list(items =  list(
                 properties.website_type = c("homepage", "facebook", "twitter", 
                                             "linkedin"),
                 properties.url = c("www.example1.com", "www.fbex1.com", 
                                    "www.twitterex1.com", "www.linkedinex1.com")))))
x[[2]] <- list(uuid = "987",
               relationships = list(websites = list(items =  list(
                 properties.website_type = c("homepage","facebook","twitter"),
                 properties.url = c("www.example2.com", "www.fbex2.com", 
                                    "www.twitterex2.com")))))
 str(x)
```

# Indexing Lists 

Lists can be subset according to their `index`, `name` or `logical statement`

```{r indexing}
# Structure
# list x = [id1 id2   ]
#     list id1 = [uuid, rel_l,] 
#         list rel = [web_l]   
#                 list web = [items_l]          
#                    list items = [char, char ]              

str(x)    
length(x) # 2
mode(x)
class(x)

# Indexing
# name the lists
names(x)
names(x) <- c("id1", "id2")
names(x)

# [ desired return value is 1st element AS a list  (keeps the box)
x1_l <- x[1]
str(x1_l)
# [[ | $ desired return value is 1st element AS element !!! (out of the box)
x1_e <- x$id1 # qui sono scesa di 1 livello!!!!
str(x1_e)
x1_e <- x[[1]] # qui sono scesa di 1 livello!!!!
str(x1_e)

# If you wanted to get the CONTENT of the 1st element
x1_e_1 <- x[[1]][[1]]
str(x1_e_1) # char "123"

x["id2"]
str(x["id2"]) # list of 1 

# indexing > 1 element (AS LISTS)
x[c("id1","id2")]       # List of 2
str(x[c("id1","id2")]) # list 
```

# Nesting/Unnesting Lists 

## Starting from 1 element/observation only ...
```{r}
# -- OPPURE --- step by step 

# get the CONTENT inside the 2nd Variable of the 1st element
x1_e_2 <-x[[1]][[2]]
str(x1_e_2) 
# list 
# $websites
# $websites$items
# $websites$items$properties.website_type
# [1] "homepage" "facebook" "twitter"  "linkedin"
# 
# $websites$items$properties.url
# [1]"www.example1.com""www.fbex1.com""www.twitterex1.com""www.linkedinex1.com"

# Make this a tibble with a list in one column  "websites"
x1_e_2_t <- as_tibble(x1_e_2)
str(x1_e_2_t) # tibble [1 × 1l-l]
# Unnest wider the column --> 2 cols:'properties.website_type' 'properties.url'
x1_e_2_t <- x1_e_2_t %>% 
  tidyr::unnest_wider( col = "websites")
x1_e_2_t #  tibble: 1 × 2(l)
# Unnestlonger the column 
# --> 4 rows/each col:"homepage""facebook""twitter""linkedin"
x1_e_2_t <- x1_e_2_t %>% 
  tidyr::unnest_longer(data =. , col = c("properties.website_type",
                                         "properties.url"  ))
x1_e_2_t # -->  tibble: 4 × 2 

# -- OPPURE --- all in one 
# turn first level list into a tibble of 1 col (which is a list of lists)
x1_e_2_t <- as_tibble(x1_e_2) %>%                   # -->  tibble [1 × 1l-l]
  # split in two cols 
  tidyr::unnest_wider(data =., col = "websites") %>% # -->  tibble: 1 × 2(l)
  # split in 4 rows 
  tidyr::unnest_longer(data =., col = c("properties.website_type",
                                        "properties.url")) # -->  tibble: 4 × 2 
str(x1_e_2_t)
```

### ... then ALL elements (faster) - {do.call}{rbind}
```{r all_r,   results='markup'}
#convert list to dataframe in long format
x_df <- do.call(rbind, lapply(x, data.frame, stringsAsFactors = FALSE))
x_df # 7 x 3

#final result
x_df_spread <- x_df %>%
  spread(relationships.websites.items.properties.website_type, relationships.websites.items.properties.url)
x_df_spread # 2 x 5 
```


### ... then ALL elements (clunkier, but clearer) - {tidyr} 
```{r all_L, results='markup'}
names(x) <- c("id1", "id2")
view(x)

# in steps 
# x_df <- x %>% as_tibble_col( .)  
# str(x_df)
# x_df <- x_df %>% unnest_wider(col = "value") # 1
# str(x_df)
# x_df <- x_df %>% unnest_longer(col = "relationships")  # 2
# str(x_df)
# x_df <- x_df %>% unnest_wider(col = "relationships") # 2
# str(x_df)
# x_df <- x_df %>% unnest_wider(col =  "items") # 2
# str(x_df)
# x_df <- x_df %>% 
# unnest_longer(col = c("properties.website_type", "properties.url"))
# str(x_df)

# --- Iterations of unnest:
x_df2 <- x %>% tibble::as_tibble_col( .) %>%  
  tidyr::unnest_wider(col ="value")  %>% 
  tidyr::unnest_longer(col ="relationships")   %>%  
  tidyr::unnest_wider(col = "relationships")  %>%  # not sure why? twice ?!?!
  tidyr::unnest_wider(col = "items")  %>%  
  tidyr::unnest_longer(col = c("properties.website_type", 
                               "properties.url")) %>% 
  # --- Lastly, group by id: 
  group_by(uuid) %>% 
  tidyr::pivot_wider(data = ., 
                     names_from = properties.website_type, 
                     values_from = c("properties.url"))

x_df2
```


### ... then ALL elements (not convinced) - {data.table}{purrr} 
```{r}
# starting point 
str(x$id1)        # 2nd level list is a  `list`  

dt_list <- purrr::map(x, as.data.table)
str(dt_list$id1) # 2nd level list is a  `data.table`  !!! 

dt <- rbindlist(dt_list, use.names =  T, fill = TRUE, idcol = F) # 2nd COLUMN is of lists  
#   uuid relationships
# 1:  123     <list[1]>
# 2:  987     <list[1]>

dt_df <- data.frame(matrix(unlist(dt), nrow=length(dt), byrow=F)) # but lost names of vars 
```


# [CASE 2] RStudio prob. _list of lists_ to df  

[RStudio response](https://community.rstudio.com/t/transform-list-of-lists-to-redundant-table/114645/7)

```{r reprex2, results='markup'}
# 1 list of 2 lists (3 & 2 obs each)
nested_l <- list(
  list( id = "a", values = c(1, 2, 3) )
  , list( id = "b", values = c(78, 99) )
)

str(nested_l)
```

## - Different ways to transform _list of lists_ into df  
```{r rstudio, results='markup'}
# 1) TRANSFORM TO DF WITH COLUMN-list
#  2x2  (1 column is a list )
df_l <- as.data.frame(do.call( rbind, nested_l))
df_l
# 2.a) TRANSFORM TO DF WITH COLUMN-vector
#  5 obs x2  (1 column is a list )
df <- tidyr::unnest(data = as.data.frame(do.call(rbind, nested_l)), 
                    cols = values)
df
# 2.b) TRANSFORM TO DF WITH COLUMN-vector
#  5 obs x2  (1 column is a list )
# `map_dfr` return a data frame created by row-binding (`map_dfc` column-binding)
df_purr <- purrr::map_dfr(.x = nested_l,
                          ~ tidyr::unnest(data.frame(.),
                                          cols = c("id", "values")))
df_purr
# 2.c) TRANSFORM TO DF WITH COLUMN-vector
# actually no need to unnest
df_purr2 <- purrr::map_dfr(.x = nested_l,
                           ~. )
df_purr2
```

# [CASE 3] RStudio prob. _list of lists (of lists)_ to df

```{r reprex 3 , results='markup'}
# HARDER ------------------------------------------------------------------
# A distinct analysis.
level3.list1 <- list(
  key1 = "name 1"
  , key2 = 14
  , key3 = 15
  , key4 = c(15, 29, 43, 57)
  , key5 = 332
  , key6 = c("A", "B", "C")
)

level3.list2 <- list(
  key1 = "name 2"
  , key2 = 28
  , key3 = 15
  , key4 = c(15, 43, 71, 99)
  , key5 = 332
  , key6 = c("Y", "Z")
)

level3.list3 <- list(
  key1 = "name 3"
  , key2 = 56
  , key3 = 22
  , key4 = c(78, 134, 190, 246)
  , key5 = 112
  , key6 = c("V")
)

# Grouped analyses.
level2.list1 <- list( level3.list1, level3.list2 )
level2.list2 <- list( level3.list3 )

# Planned samplings.
level1.list1 <- list(
  code = "ABC123"
  , type = "spot"
  , matrix = "water"
  , analyses = level2.list1
)

level1.list2 <- list(
  code = "GHS332"
  , type = "mix"
  , matrix = "water"
  , analyses = level2.list2
)

# All the samplings.
level0.list <- list(level1.list1, level1.list2)

str(level0.list)
```


## - Different ways to transform list of lists (of lists) into df  

Using `purrr::map_dfr` returns a data frame created by row-binding (last column is of lists)

```{r}
# "Flattening" 
# 3x4 [l]
df_l.lev2 <- purrr::map_dfr(level0.list, ~.)
```

Using `tibble` + `unnest_...` returns a data frame only of atomic vectors

```{r}
# 1) Step by step 
u1 <- level0.list %>% tibble() %>%
  unnest_wider(col = ".")
u2 <- u1 %>% unnest_longer(col = "analyses")
u3 <- u2 %>% unnest_wider(col = "analyses")
u4 <- u3 %>% unnest_longer(col = "key4")
u5 <- u4 %>% unnest_longer(col = "key6")

# 2) All in one with piping
u <- tibble(level0.list) %>%
  # integers are just a quick way to select the columns to operate on
  unnest_wider(1)   %>%  # u1
  unnest_longer(4)  %>% # u2
  unnest_wider(4)   %>%  # u3
  unnest_longer(7)  %>% # u4
  unnest_longer(9)     # u  24 obs X 9 var
```


