---
title: "Github authentication with PAT"
# description: |
#   A short description of the post.
#author: []
date: 2022-05-28
output:
  distill::distill_article:
    self_contained: false
    toc: TRUE
    toc_depth: 2
    toc_float: TRUE
draft: true
categories: [tools]
editor_options:
  markdown:
    wrap: 100
---
 
  
# Setting up GitHub & authentication 

## Reference / tutorials
+ [Tutorial on R & Github setup!!!](https://rworkshop.uni.lu/install_tutorial.html)
+ [Setting Up a Personal Access Token via R](https://carpentries.github.io/sandpaper-docs/github-pat.html) 
<!-- + [Multiple SSH](https://www.section.io/engineering-education/using-multiple-ssh-keys-for-multiple-github-accounts/)
+ [Github from psw -> SSH](https://community.rstudio.com/t/github-rstudio-how-to-switch-from-password-to-ssh/97096/2) -->

 ```{r, echo=FALSE}
if (!require("pacman")) {
	install.packages('pacman' )
}
# pacman::p_load_gh("mitchelloharawild/icons")
pacman::p_load(usethis,# Automate Package and Project Setup
 	      gitcreds, # Query and set git credentials
            credentials,
            #gh, # Query the GitHub API
            #gert,
            fontawesome) 
  
 ```


## Verify your git installation & configuration (from terminal)

[see info on Atlassian](https://www.atlassian.com/git/tutorials/setting-up-a-repository/git-config)
The `git config` command is a convenience function that is used to set Git configuration values on a global or local project level. These configuration levels correspond to `.gitconfig` text files. 

> **LEVELS of git config** 

1. By default, git config will write to a **local level** if no configuration option is passed. Local level configuration is _applied to the context repository_ git config gets invoked in. Local configuration values are stored in a file that can be found in the repo's .git directory: `.git/config`. 

`git config --local`

2. **Global level** configuration is user-specific, meaning it is _applied to an operating system user_. Global configuration values are stored in a file that is located in a user's home directory. `~ /.gitconfig` on unix systems and `C:\Users\\.gitconfig` on windows.

`git config --global`

3.  **System level** System-level configuration is applied across an entire machine. This _covers all users on an operating system and all repos_. The system level configuration file lives in a gitconfig file off the system root path. `$(prefix)/etc/gitconfig` on unix systems. 

`git config --system`


```{bash, eval=FALSE}
# check version
git version
```

> Since these are hidden files I need `ls -al` to see them 

```{bash, eval=FALSE}
# see hidden config files
cd ~
ls -al .git*

# or using git command...   
git config --list --show-origin

# check configuration 
git config --list

# among other stuff
# ...
# credential.helper=osxkeychain
# user.name=Lulliter
# user.email=lmm76@georgetown.edu
# aliases!!!!! 

# ask child property like  
git config user.email
# lmm76@georgetown.edu

# write a child property like  
git config --global user.email "lmm76@georgetown.edu"
```

## If needed, introduce yourself to git (from R)
```{r, eval=FALSE}
# {NOT RUN}
## set your user name and email:
library(usethis)

# ---STOP!---- not sure this still exists 
if (FALSE) {
# set the user's global user.name and user.email
use_git_config(user.name = "Lulliter", user.email = "lmm76@georgetown.edu")

# (OR set the user.name and user.email locally, i.e. for current repo/project
use_git_config(
  scope = "project",
  user.name = "Lulliter", 
  user.email = "lmm76@georgetown.edu"
)
}
```


# 3 ways to authenticate in Github

0. ~~password authentication~~ [deprecated since August 2021]
1. PATs (Personal Access Tokens) Authentication [*me, | recommended beginners]
2. SSH Key Authentication
3. Applications API tokens

> _"if you have no burning reason to pick SSH"_ (1) is [recommended by Jenny Bryan](https://happygitwithr.com/https-pat.html) 

## 3.1.a How I configured Github PAT  
Clear instructions [here](https://betterprogramming.pub/dealing-with-github-password-authentication-deprecation-1b59ced90065),  and [here (JB)](https://happygitwithr.com/https-pat.html), and [here (ROpensci)](https://ropensci.org/blog/2020/07/07/github-pat/).

A proper PAT is characterized by the following four features:

+ **Unique**: It is distinctive to GitHub and can be generated per device or usage.
+ **Revocable**: It can be individually revoked at any time without needing to update any credentials.
+ **Limited**: It is narrowly scoped by definition and provides access only to a limited set of operations.
+ **Random**: Unlike passwords, it is not subject to dictionary or brute force attacks.

> With the PAT, I will be performing Git operations over HTTPS protocol (as opposed to SSH). 

#### 1) Create Github PAT 

Either: 

a. Go to https://github.com/settings/tokens and click “Generate token”.
b. Create a personal access token for authentication (from R) with: 

```{r ,eval=FALSE}
usethis::create_github_token() 
## in case usethis version < 2.0.0: usethis::browse_github_token() 
# (or even better: update usethis!)
```

#### 2) Provide this PAT next time a Git operation asks for your password
Either: 

a. Put the PAT into the Git credential store
b. Use `gitcreds::gitcreds_set()` (from R) --> this will prompt to paste the PAT 

#### 3) Managing/Verifying git credential stored 

> Where do I see it on Github?
`r fa(name = "github")` --> **settings** --> **Developer settings** --> Personal Access Tokens   

> WARNING: `credentials::git_credential_ask("https://github.com")` actually exposes the PAT password

```{r ,eval=FALSE}
credentials::credential_helper_get()
      # [1] "osxkeychain"

# to see all of them 
credentials::git_credential_ask("https://github.com")
      # $protocol
      # [1] "https"
      # $host
      # [1] "github.com"
      # $username
      # [1] "PersonalAccessToken"
      # $password
      # [1] "ghp_...8tNB" !!!!!!!!!!!!!!!!!!!!!!!!!!!!!!
      # attr(,"class")
      # [1] "git_credential"

# --- other useful things to check 
#usethis::gh_token_help()
usethis::git_sitrep()
gh::gh_whoami()
# {
#   "name": "Luisa Mimmi",
#   "login": "Lulliter",
#   "html_url": "https://github.com/Lulliter",
#   "scopes": "gist, repo, user, workflow",
#   "token": "ghp_...52eR"
# } 
```

--- 
## 3.3 What are SSH keys?
Secure Shell Protocol (SSH) is a cryptographic network protocol that allows a single computer to connect with a server over the internet securely. SSH is best used for accessing remote servers.

## Check which ssh I already have 

> I have a github SSH but I am not using it yet!

```{bash}
 ls -al ~/.ssh/   

#  total 40
# drwx------   7 luisamimmi  staff   224 Sep  8  2017 .
# drwxr-xr-x@ 99 luisamimmi  staff  3168 May 26 19:50 ..
# -rw-------@  1 luisamimmi  staff  1766 Jun  6  2015 github_rsa
# -rw-r--r--   1 luisamimmi  staff   401 Jun  6  2015 github_rsa.pub
# -rw-------   1 luisamimmi  staff  3243 Feb 21  2018 id_rsa
# -rw-r--r--@  1 luisamimmi  staff   746 Feb 21  2018 id_rsa.pub
# -rw-r--r--@  1 luisamimmi  staff  2783 Sep  8  2017 known_hosts
```


## Which one am I currently using for my [Github account](https://github.com/Lulliter)?
errrr **NONE!**  

`r fa(name = "github")` --> **settings** --> **SSH keys** : You don't have any public SSH keys in your GitHub account!!!!

Although it does appear to be there as told by `library(credentials)`  

```{r}
library(credentials)
#  Found git version 2.15.1
# Supported HTTPS credential helpers: cache, cache--daemon, store, osxkeychain

# 'git
# Found OpenSSH_8.1p1, LibreSSL 2.7.3
# Default SSH key: /Users/luisamimmi/.ssh/id_rsa
```

 
 ...tbc 
